### 2024-10-02 manuscript/supplement.qmd
initial commit

---
title: "Supplementary Material"
subtitle: "..."

echo: false
number-sections: true

format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    fig-cap-location: top
    crossref:
      fig-title: '**Supplementary Figure**'
      tbl-title: '**Supplementary Table**'

editor: visual
bibliography: references.bib
#csl: american-journal-of-epidemiology.csl
---

```{r}
#| label: setup_supplement
#| include: false

# table_cap <- captioner(prefix = "Supplementary Table")
```

**Authors**: ...

$\vspace{0.25cm}$

*Supplementary Material last updated: `r Sys.Date()`*

{{< pagebreak >}}

-->

# Supplementary Tables

...

{{< pagebreak >}}

# Supplementary Figures

...

{{< pagebreak >}}

# References {.unnumbered}

::: refs
:::