### 2024-12-02 manuscript/supplement.qmd
not on nsclc trials

---
title: "Supplementary Material"
subtitle: "A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)"

echo: false
number-sections: true
toc: true

format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    fig-cap-location: top
    crossref:
      fig-title: '**Supplementary Figure**'
      tbl-title: '**Supplementary Table**'

editor: visual
bibliography: references.bib
#csl: american-journal-of-epidemiology.csl
---

```{r}
#| label: setup_supplement
#| include: false

library(here)
library(knitr)

# table_cap <- captioner(prefix = "Supplementary Table")
```
{{< pagebreak >}}

# Supplementary Figures

```{r}
#| label: fig-consort-nsclc
#| fig-cap: "CONSORT diagram for non-small cell lung cancer (NSCLC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_nsclc.png"))
```

*Note that @fig-consort-nsclc results in four shortlisted candidates because both CheckMate017/057 have been both shortlisted and only differ in the squamous versus nonsquamous histological eligibility of the trial population.*

{{< pagebreak >}}

```{r}
#| label: fig-consort-breast
#| fig-cap: "CONSORT diagram for breast cancer (BC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_bc.png"))
```

{{< pagebreak >}}

```{r}
#| label: fig-consort-crc
#| fig-cap: "CONSORT diagram for colorectal cancer (CRC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_crc.png"))
```

{{< pagebreak >}}

```{r}
#| label: fig-consort-mm
#| fig-cap: "CONSORT diagram for multiple myeloma (MM) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_mm.png"))
```

{{< pagebreak >}}

# Supplementary Tables

{{< pagebreak >}}

$\vspace{0.25cm}$

*Supplementary Material last updated: `r Sys.Date()`*

{{< pagebreak >}}

# References {.unnumbered}

::: refs
:::