### 2025-03-04 manuscript/manuscript.tex
Manuscript updated

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}

\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading,figureheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Process for Emulating Comparative Oncology Trials with Real-world Evidence Studies},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{A Process for Emulating Comparative Oncology Trials with
Real-world Evidence Studies}
\author{}
\date{}

\begin{document}
\maketitle

\textbf{Authors}: Janick Weberpals\textsuperscript{1}, Kenneth L.
Kehl\textsuperscript{2}, Donna R. Rivera\textsuperscript{3}, Pallavi
Mishra-Kalyani\textsuperscript{3}, Georg Hahn\textsuperscript{1},
Priyanka Anand\textsuperscript{1}, Yanina Natanzon\textsuperscript{4},
Andrew J. Belli\textsuperscript{5}, Ching-Kun Wang\textsuperscript{5},
Jenna Collins\textsuperscript{6}\textbf{,}
JonathanKish\textsuperscript{6}\textbf{,} Janet
Espirito\textsuperscript{7}, Nicholas J Robert\textsuperscript{7},
Robert J. Glynn\textsuperscript{1}, Sebastian
Schneeweiss\textsuperscript{1}, Shirley V. Wang\textsuperscript{1}

\ul{Author affiliations:}

\textsuperscript{1} Division of Pharmacoepidemiology and
Pharmacoeconomics, Department of Medicine, Brigham and Women's Hospital,
Harvard Medical School, Boston, MA, USA

\textsuperscript{2} Dana-Farber Cancer Institute, Boston, MA, USA

\textsuperscript{3} Oncology Center of Excellence, US Food and Drug
Administration, Silver Spring, MD, USA

\textsuperscript{4} ConcertAI, Cambridge, MA, USA

\textsuperscript{5} COTA, Inc., New York, NY, USA

\textsuperscript{6} Flatiron Health, Inc., New York, NY, USA

\textsuperscript{7} Ontada, Boston, MA, USA

\ul{\textbf{Correspondence:}}

Shirley V. Wang, PhD

Division of Pharmacoepidemiology and Pharmacoeconomics,

Department of Medicine, Brigham and Women's Hospital, Harvard Medical
School,

1620 Tremont Street, Suite 3030-R, Boston, MA 02120, USA

Phone: +1 617-278-0932

Fax: + 1 617-232-8602

Email: \url{SWANG1@BWH.HARVARD.EDU}

\ul{\textbf{Article type:}} Review

\ul{\textbf{Manuscript word count:}} 4,131 words / 8,000 words

\ul{\textbf{Abstract word count:}} 248 words / 250 words

\ul{\textbf{Tables:}} 3

\ul{\textbf{Figures:}} 3

\ul{\textbf{Supplementary material:}} Supplementary figures and material

\ul{\textbf{Short running title}}: Emulation of Comparative Oncology
Trials with Real-world Evidence (ENCORE)

\ul{\textbf{Keywords:}} Oncology, Real-World Evidence, Trial emulation,
EHR

\ul{\textbf{Funding Statement:}} This project was supported by an FDA
BAA contract with contract number 75F40122C00181.

\ul{\textbf{Competing Interests Statement:}} Dr.~Weberpals is now an
employee of AstraZeneca and owns stocks in AstraZeneca. Dr.~Kehl has
received research funding from Meta, Inc.~to his institution. Drs.
Espirito and Robert are employees of McKesson and own McKesson stock.
Dr.~Wang has consulted ad hoc for Exponent Inc.~and MITRE a federally
funded research center for the Centers for Medicare and Medicaid
Services on unrelated work. Dr.~Schneeweiss is participating in
investigator-initiated grants to the Brigham and Women's Hospital from
Bayer and UCB unrelated to the topic of this study. He consults for and
owns equity in Aetion Inc., a software manufacturer. He is an advisor to
Temedica GmbH, a patient-oriented data generation company. His interests
were declared, reviewed, and approved by the Brigham and Women's
Hospital in accordance with their institutional compliance policies.

\ul{\textbf{Data sharing statement:}} No data was analyzed as part of
this project.

\ul{\textbf{Analytic code sharing statement:}} Code to reproduce this
manuscript, including all figures and tables, can be found at
\url{https://github.com/janickweberpals/encore-process-manuscript}.

\ul{\textbf{\emph{Proposed target journals:}}} \emph{CPT =\textgreater{}
JCO CCI =\textgreater{} \ldots{}}

\emph{Manuscript last updated: 2025-03-04 18:43:42.372054}

\newpage{}

\section*{Abstract}\label{abstract}
\addcontentsline{toc}{section}{Abstract}

Real-world evidence (RWE) studies are increasingly used to complement
evidence from randomized controlled trials (RCTs), contextualizing the
effectiveness and safety of medical interventions as delivered in
clinical practice. Advancements in the curation and accessibility of
electronic health record data (EHR) have presented the opportunity to
investigate disease domains such as oncology, where administrative
healthcare claims databases alone are not fit-for-purpose. The RCT
DUPLICATE initiative has previously enhanced understanding of when RWE
studies come to the same causal conclusions by emulating trials in
non-oncology indications. Here, we present the Emulation of Comparative
Oncology Trials with Real-world Evidence (ENCORE) project, which aims to
extend this work to oncology. ENCORE will emulate 12 RCTs in four
oncology-specialized EHR databases across four different cancer
indications, specifically non-small cell lung cancer, breast cancer,
colorectal cancer, and multiple myeloma. It will place a special
emphasis on systematic evaluation of fitness of data in relation to the
study design and statistical analysis for a particular research
question, and pre-registration of study protocols prior to analysis.
Agreement of treatment effect estimates between RCTs and their
respective emulations will be assessed using pre-specified criteria.
Through extensive sensitivity analyses benchmarked against RCT results,
the ENCORE project will inform understanding of how measurement, design,
and analytic decisions influence the validity of oncological RWE
studies.

\newpage{}

\section{Background}\label{background}

Randomized controlled trials (RCTs) have been the go-to methodology for
establishing the efficacy and safety of medical products. Under the
21\textsuperscript{st} Century Cures Act directive,\textsuperscript{1}
the Food and Drug Administration (FDA) established a framework to
increasingly consider real-world evidence (RWE) generated from
routine-care health data such as electronic health records (EHR) to
evaluate and contextualize the comparative safety and effectiveness of
novel cancer therapies.\textsuperscript{2} Accounting for 21\% of all
approvals, oncology was the disease area with the most FDA drug
approvals in 2023.\textsuperscript{3} RWE has particularly important
potential to complement evidence coming from RCTs in the field of
precision oncology, where potential use cases include the assessment of
effectiveness in specific patient populations that are underrepresented
in RCTs, the construction of external control arms in single-arm trials
where active recruitment to a randomized trial may not be feasible, or
the discovery of biomarkers among pan-tumor populations that harbor
specific genomic and immuno-pathological signatures.

However, the validity and transportability of results derived between
RWE studies and RCTs depends on multiple factors. Frequently referenced
limitations include, lack of baseline randomization, missing data, small
study sizes, data discontinuity,\textsuperscript{4, 5} adoption of
changes in guidelines in real-world care and the inability to measure
and emulate common eligibility criteria, including prognostic factors,
and standardized response assessments in real-world data
(RWD).\textsuperscript{6} While there are already published examples of
oncology trial emulations,\textsuperscript{6--8} a systematic and scaled
approach to emulate a diverse set of different oncology trials with
multiple heterogeneous databases is necessary to gain confidence in the
validity of RWE studies and to provide context as to which questions can
be validly answered.

The RCT DUPLICATE initiative\textsuperscript{9} increased our
understanding of when RWE studies can come to causal conclusions on
treatment effects by benchmarking results against RCTs under the
assumption that each RCT finding reflects a causal treatment effect. In
settings where the RCT designs could be emulated well, RWE studies came
to the same conclusions.\textsuperscript{10} However, prior work from
RCT-DUPLICATE has focused primarily on emulating trials in non-oncology
settings using claims databases.

The \emph{Emulation of Comparative Oncology Trials with Real-world
Evidence} (ENCORE) project\textsuperscript{11} aims to extend this work
to oncology. Studies in oncology come with their own unique set of
challenges which must be systematically explored and understood.
Building on a process co-developed with the FDA through RCT
DUPLICATE,\textsuperscript{9} this expansion to oncology will emulate 12
randomized clinical trials using multiple specialty oncology EHR data
sources. The process will emphasize transparency and include documented
data fitness assessment\textsuperscript{12, 13} for each RWD source with
respect to each trial emulation as well as extensive sensitivity
analyses to assess robustness of findings.

The objectives of this project are to 1) develop state-of-the-art
methodological approaches and 2) to apply them to create insights that
may provide guidance on the potential use of RWE for regulatory science
in oncology. To achieve these objectives, this demonstration project
will systematically emulate 12 oncology trials across four cancers and
assess the agreement of treatment effect estimates between RCTs and
their respective emulations.

Here, we describe the design and process for the selection of the 12
oncology RCTs, the assessment of the database quality and selection,
protocol development and pre-registration, study design and statistical
analysis, and pre-specified agreement metrics to evaluate the
concordance between RCTs and emulations.

\section{Methods}\label{methods}

A visual summary of the entire systematic process from trial selection
to final results is provided in Figure~\ref{fig-process}.

\subsection{Trial selection}\label{trial-selection}

The focus of ENCORE is to maximize potential learnings on when RWE
studies can or cannot yield similar results compared to RCTs. To that
end, the emphasis of the project is on trials of therapies for the most
common cancers for which there has been substantial therapeutic
development in recent years. After review and discussion with clinical
and regulatory experts, four cancer indications were identified:
non-small cell lung cancer, breast cancer, colorectal cancer and
multiple myeloma. For each cancer type, we aim to conduct three trial
emulations using multiple databases accessible for the scope of this
project (i.e., the total number of emulations will equal 12 trials x
\emph{n} databases which are found fit-for-purpose for each trial).

We used a semi-automated process for trial selection where the
eligibility criteria are documented in CONSORT diagrams showing reasons
for excluding RCTs for each cancer type. The search was conducted using
the \emph{Aggregated Analysis of ClinicalTrials.gov} (AACT) database
which is a publicly available relational database developed and
maintained by the Clinical Trials Transformation Initiative (CTTI) that
contains all information (protocol and result data elements) about every
study registered on ClinicalTrials.gov.\textsuperscript{14} To identify
eligible trials, we used a combined search query strategy of the
National Library of Medicine (NLM)-controlled \emph{Medical Subject
Headings \textbf{(}}MeSH) term and a free keyword search for the
respective cancer indication in the \emph{conditions}, \emph{studies}
and \emph{detailed\_descriptions} fields of each trial entry on
ClinicalTrials.gov.

Eligible trials had to fulfill the following basic criteria:

\begin{itemize}
\item
  Interventional
\item
  Randomized
\item
  Intervention model: parallel assignment
\item
  Industry-sponsored
\item
  Trial start in 2011 or later
\item
  Primary purpose was to study treatment effects
\item
  Overall survival must be one of the endpoints reported (either as
  hazard ratio or median overall survival time)
\item
  Recruitment status: `Completed' or `Active, not recruiting'
\item
  Feasibility and clinical relevance (latter was defined as treatment or
  paradigm-changing trials or such that challenged existing treatment
  policies)
\end{itemize}

The rationale and operationalization of each criterion is listed in
detail in Table~\ref{tbl-criteria}. We will mainly consider pivotal
interventional, randomized trials after 2011 because treatment
guidelines among included cancer indications have undergone significant
changes in recent years. Due to the rapid adoption of new breakthrough
therapies in routine care, it is unlikely to find patients who may be
still treated with outdated treatment regimens in current clinical
practice. Conversely, we excluded trials with results published too
recently. To allow for enough data and follow-up time accrual in the
databases used for this project. Although there have been substantial
methodological advancements to increase our understanding on the
emulation and comparison of real-world progression-free survival (PFS)
and objective response rates (ORR) to a
RECISTv1.1\textsuperscript{15}-based PFS and ORR assessment in
RCTs\textsuperscript{16, 17}, imaging-based evaluations still hold a
level of granularity which may not be necessarily reflected in
chart-abstracted assessments of a patient's progression in routine
care.\textsuperscript{18, 19} The timing and cadence of intervals
between progression assessments can differ between RCTs and routine care
which may result in measurement error and bias.\textsuperscript{20}
Given this, and the large number of other methodological challenges like
missing data, small sample sizes, data discontinuity and rapidly
changing guideline treatments, we focus on the emulation of overall
survival (OS) as the endpoint of interest. Therefore, we include only
trials that have reported overall survival (OS) as one of the
pre-specified endpoints in the protocol.

While most trial eligibility criteria could be operationalized in an
automated fashion, our final criterion was an assessment of emulation
feasibility and clinical relevance. This criterion involved extensive
human review. The critical points considered in this step include an
initial feasibility assessment of the data fitness,\textsuperscript{19}
including assessment of whether critical eligibility criteria (e.g.,
biomarker status) and prognostic factors (e.g., ECOG performance score)
are measurable and whether preliminary study size counts are reasonable.
Lastly, trial candidates were ranked and shortlisted into primary and
runner-up candidates based on their clinical and regulatory relevance.

A list of tentative, shortlisted primary candidates is presented in
Table~\ref{tbl-rcts} and the corresponding selection process is
illustrated in the CONSORT diagrams (Supplementary Figures 1-4).
Naturally, the majority of trials cover advanced (locally advanced,
inoperable, recurrent/progressive disease) or metastatic cancer
populations because a large proportion of drug development efforts start
in these settings. A key objective that we aim to foster with the
shortlisted trials is to achieve a better understanding how different
disease settings (early, late), line settings ({[}neo{]}adjuvant, first
line, advanced lines of therapy), therapy protocols (monotherapy,
combination therapy) and population characteristics (simple versus
complex genetic or immunological signatures) can be emulated using RWD.
If more thorough feasibility assessments suggest that the threat of bias
from mis-measurement of key study parameters or residual confounding
remains high, or that the study size is not sufficient, then runner-up
candidates will be considered instead.

\subsection{Databases}\label{databases}

The ENCORE project will use data from four oncology-specific electronic
health records (EHR)-derived data sources (in alphabetical order):
ConcertAI, COTA, Flatiron Health, and Ontada/McKesson. All available
databases draw from a comprehensive national sample of patients with
cancer in the US with the detailed EHR-based data necessary to study
medication effectiveness in oncology. A detailed description and
methodology on how patients are sampled will be provided with each trial
emulation protocol. For ENCORE, not all databases will be available for
each cancer indication and the names of the databases will be blinded
and referred to as ENCORE DataBase (EDB) 1, 2, 3 and 4 for the final
reporting of results (the numbering does not coincide with the above
order of mention of the databases). If more than one database is
considered fit-for-purpose for a respective trial emulation, the best
possible analytic model will be employed for each database separately
and final treatment effect estimates will be pooled using a
meta-analytic approach using fixed effects (weights reflect the inverse
of the variance of each database study) and random effects (weights
reflect the inverse of the variance of each database study plus an
additional variance term that quantifies the assumed heterogeneity
between databases) models.\textsuperscript{21}

\subsection{Protocol development}\label{protocol-development}

For each shortlisted and selected RCT, a detailed protocol,
pre-specifying key elements of the trial emulation, will be developed
using the HARPER protocol template\textsuperscript{22} recommended for
regulatory submissions of RWE studies\textsuperscript{23} and will be
registered on ClinicalTrials.gov after review by a clinical and FDA
regulatory expert panel. Following the target trial emulation framework,
we will provide an explicit statement and rationale on how each element
will be emulated including database selection, covariate measurement,
operationalization of key eligibility criteria, study design, data
analysis and causal contrasts of interest.\textsuperscript{24, 25} Since
it is common that oncology RCTs update OS estimates periodically based
on accrued follow-up time, the protocol will give a brief summary of
each emulated RCT and specify which target OS estimates will be used to
compare agreement metrics to (see Section~\ref{sec-agreement-metrics}).
All eligibility criteria will be extracted based on publications,
publicly available protocols and statistical analysis plans of the
selected RCT.

\subsubsection{Emulation feasibility}\label{emulation-feasibility}

\paragraph{\texorpdfstring{\textbf{Fit-for-purpose
data}}{Fit-for-purpose data}}\label{fit-for-purpose-data}

Real-world data fitness and emulation feasibility for each shortlisted
candidate trial will be assessed in multiple steps based on guidance of
the oncology quality, characterization, and assessment of real-world
data (Oncology QCARD) Initiative.\textsuperscript{19} The first step
assesses if relevant variables like exposure/line of therapy, outcomes,
and covariates are generally available, measured and operationalizable
in routine-care. Since many oncological RCTs in recent years have
focused on selected, biomarker-defined populations, subtleties in
measurement and operationalizability of specific biomarkers must be
reflected to ensure a representative and large enough study population.
For example, immune checkpoint inhibitors have significantly changed the
cancer treatment landscape since the first approvals of the CTLA-4
inhibitor ipilimumab in 2011 and the PD-1 inhibitors pembrolizumab and
nivolumab in 2014 in the United States. With many trials that have
followed thereafter, the operationalization of the expression of the
PD-L1 biomarker in RCTs (e.g., as a percent staining, tumor proportion
score or combined positive score) has also evolved, and PD-L1
'\emph{positivity'} may have different definitions across calendar years
based on different cut-off values.

According to the \emph{Structured Process to Identify Fit-For-Purpose
Data} (SPIFD) framework\textsuperscript{13}, the next step will outline
tables that describe how eligibility criteria will be ascertained using
a color-coded heatmap that will indicate the level of confidence on how
well each criterion can be emulated in each selected database. As there
are general eligibility criteria in oncology trials which either will
not be possible to emulate (e.g., physician-assessed survival prognosis
of x months) or that are clinically not relevant for the emulation of
the trial (e.g., male patients should be willing to use barrier
contraception), the study team will decide on key eligibility criteria
for the emulation of the trial.

We will additionally provide a definition on how exposures, outcomes and
covariates are exactly defined and operationalized in each respective
database. There will be special emphasis on how exposure, in context of
their respective disease and line of therapy settings, and the OS
outcome will be emulated. For all considered databases, the OS endpoint
is typically a composite that, depending on the underlying database, can
be derived from different sources comprising EHR documentation, social
security death index, obituary and other linkages. Given that not all
relevant sources that provide mortality data are synchronized and
updated uniformly, sensitivity analyses with more conservative (i.e.,
earlier) censoring dates will be considered for each trial emulation to
mitigate the potential impact of ghost-time bias.\textsuperscript{26}

\paragraph{\texorpdfstring{\textbf{Descriptives and data
exploration}}{Descriptives and data exploration}}\label{descriptives-and-data-exploration}

Critical aspects when emulating oncology trials are the choice and
estimation of the appropriate estimand of interest.\textsuperscript{27}
Particularly when emulating pivotal trials of paradigm-changing
treatments, multiple aspects need to be considered such as the
contemporaneity of the (historical) control cohort, the adoption rate of
the novel intervention in routine care, the magnitude of the clinical
treatment benefit and the rate in which (particularly patients in the
control arm) discontinue or cross-over to the interventional treatment,
which could lastly bias emulated treatment effects towards the null. To
that end, comprehensive data explorations will be performed as part of
the protocol development to contextualize these parameters and (if
reported) draw comparisons to the emulated trial. Examples for such
standard diagnostics are visualized in Figure~\ref{fig-initiators}. All
exploratory analyses will be conducted blinded towards the outcome to
avoid influencing study design and analytic choices.

The distribution of patient characteristics, stratified by exposure
status, will be examined in Table 1's before and after applying
eligibility criteria and contrasted with the distributions of patient
characteristics of the original RCT. Initial propensity score matching
or weighting methods will be applied to ensure that measured
pre-exposure covariates can be balanced, exposure cohorts are
conditionally exchangeable at baseline and resulting sample sizes are
still sufficient after matching or weighting. At this stage, all
exploratory analyses will be conducted blinded towards the outcome to
not bias any study design and analytic choices based on known outcome
information.

\paragraph{\texorpdfstring{\textbf{Statistical power
considerations}}{Statistical power considerations}}\label{statistical-power-considerations}

Causal analyses of observational data may not have the same
pre-requisites in terms of formal hypothesis testing and statistical
power as RCTs since the number of `recruited' patients is given and
cannot be influenced.\textsuperscript{28} For this project, however, the
emulations in EHR data must have at least equal power to the relevant
trial for interpretation of the pre-specified agreements metrics between
RCT and RWD results. Since the main outcome of interest is defined as
time to all-cause mortality (OS), the estimation of the statistical
power is driven by the number of events rather than the number of
patients. To assess if the overall number of events, unstratified by
exposure, is sufficient such that a significant difference can be
detected based on the original RCT-reported hazard ratio (HR), the
statistical power (1-\(\beta\)) will be estimated using Schoenfeld's
sample-size formula for the proportional-hazards regression
model.\textsuperscript{29}

\subsection{Agreement metrics}\label{sec-agreement-metrics}

To formally compare treatment effects between RCTs and their respective
emulations, we will adapt the approach of the RCT-DUPLICATE
project.\textsuperscript{9, 30} That is, for the primary endpoint of
interest (HR for OS and corresponding 95\% confidence intervals), we
will derive three qualitative agreement metrics: statistical agreement,
estimate agreement and agreement based on the standardized mean
difference (SMD). Examples are illustrated Table~\ref{tbl-metrics}.

\textbf{Statistical significance agreement}: agreement between RCT and
emulated trial treatment effect with regards to directionality and
statistical significance.

\textbf{Estimate agreement}: agreement that the estimated RWE treatment
effect is within the 95\% CI of the RCT treatment effect estimate.
Provided that for some emulations, the power of the RWE study may be
larger than that of the original RCT, this could lead to situations
where there is no statistical significance agreement although the
treatment effect estimates are highly overlapping but with the RCT
estimate crossing the null (or vice versa in case the RCT has a larger
power than the RWE emulation).

\textbf{SMD agreement}: quantification of the agreement between the
emulated RWE and RCT treatment effect estimate. The SMD is calculated as

\[
SMD = \frac{\hat{\theta}_{RCT} - \hat{\theta}_{RWE}}{\sqrt{\text{Var}(\hat{\theta}_{RCT}) + \text{Var}(\hat{\theta}_{RWE})}}
\]

where \(\hat{\theta}_{RCT}\) and \(\hat{\theta}_{RWE}\) are the
treatment effect estimates (hazard ratios or median survival time
differences) and \(\text{Var}(\hat{\theta}_{RCT})\) and
\(\text{Var}(\hat{\theta}_{RWE})\) are the corresponding variances for
RCT and RWE, respectively. The resulting SMDs will be interpreted such
that with an SMD of 1.00, the effect estimate from the RCT and the RWE
emulation are 1 standard deviation apart. For an \(\alpha\)-level of
0.05, the null hypothesis of no difference would be rejected whenever
\(|Z| > 1.96\).

For the secondary endpoints of interest (e.g., median survival time or
survival probabilities) only the SMD agreement metric will be
applicable.

\subsection{Study design and statistical
analysis}\label{study-design-and-statistical-analysis}

The study design for each trial emulation will be visualized as part of
the protocol using a graphical depiction of the exact measurement
windows of eligibility criteria, washout periods and covariates relative
to the cohort entry time.\textsuperscript{31}

\subsubsection{Missing data}\label{missing-data}

To establish an analytic cohort, key eligibility criteria will be
applied in which patients with missing values in eligibility criteria
are considered eligible in the respective attrition steps to allow for
thorough missing data investigations. These missing data investigations
will empirically assess assumptions on potential underlying missingness
mechanisms according to Rubin's classification of missing data (i.e.,
missing completely at random {[}MCAR{]}, missing at random {[}MAR{]} and
missing not at random {[}MNAR{]}).\textsuperscript{32} To that end, we
will adopt a principled process on missing data that empirically
evaluates different aspects across partially observed covariates based
on three group diagnostics.\textsuperscript{33, 34} These diagnostics
cover (1) comparisons of patients characteristics with and without an
observed level of the partially observed covariate, (2) ability to
predict missingness given observed data, and (3) assessments if outcomes
between patients with a missing value are systematically different.
Together with expert domain knowledge and assumptions about the
underlying missing data structure through canonical causal
diagrams,\textsuperscript{35} this will inform decisions regarding the
in- or exclusion of patients with missing values in key eligibility
criteria and potential sensitivity analyses to assess the robustness of
these decisions.\textsuperscript{36}

While the MAR assumption is a strong assumption to hold across all
considered covariates, it was shown that especially in the context of
partially observed covariate data (as opposed to missing exposure and
outcome data), only mechanisms in which a covariate causes its own
missingness leads to critical bias (MNAR).\textsuperscript{35} Hence,
methodologies which retain patients and give the potential to adjust for
a broader set of prognostic factors (e.g., multiple
imputation\textsuperscript{37} or doubly robust
methods\textsuperscript{38}) may be preferred over complete case
analyses.

\subsubsection{Endpoints and propensity score
analyses}\label{endpoints-and-propensity-score-analyses}

Due to its ubiquity in oncology trials, the primary parameter of
interest in ENCORE will be defined as the marginal hazard ratio (HR)
coefficient for the treatment comparison for time to all-cause mortality
(OS).\textsuperscript{39} We will also consider alternative endpoints on
an absolute risk scale such as median survival times, survival
probabilities at pre-defined time points during
follow-up\textsuperscript{40} and restricted mean survival times as
secondary endpoints of interest.

For the estimation of marginal treatment effects, we will employ
propensity score methods to adjust for measured confounding between
treatment arms. The selection of important prognostic covariates will be
based on expert clinical knowledge and published literature on
prognostic scores in oncology.\textsuperscript{41} The implementation of
propensity scores in combination with multiple imputation will follow
the `\emph{within}' methodology as described by Leyrat et
al.\textsuperscript{42, 43} That is, propensity score matching or
weighting will be applied to each imputed dataset. The marginal
treatment effect will then be estimated in each imputed and matched or
weighted dataset separately and pooled into a final estimate following
Rubin's rule.\textsuperscript{44, 45} This approach has been shown to
lead to unbiased estimates across different simulated scenarios with a
sufficient estimation of the variance.\textsuperscript{42}

To asses the balance of pre-exposure covariates after matching or
weighting on each imputed dataset, the average SMD and corresponding
minimum and maximum SMD range will be visualized (see example in
Figure~\ref{fig-balance}). Covariate balance is typically considered at
a SMD \textless{} 0.1.\textsuperscript{46} Further, we will compute the
average post-matching or post-weighting
C-statistics.\textsuperscript{47} In addition, we will use a published
prognostic score for OS\textsuperscript{41} as a balance measure to
visually assess if the prognostic score is balanced between treatment
arms after propensity score matching or weighting as this approach was
described to show the highest correlations with bias compared with other
balance measures and not affected by model
misspecification.\textsuperscript{48}

Similarly, survival probabilities for individual time points will be
estimated in each imputed and propensity score matched or weighted
dataset according to the Kaplan-Meier method.\textsuperscript{40} Since
survival probabilities typically do not follow normal distributions
which are required to apply Rubin's rule, these will be transformed
through a complementary log-log transformation \(log(-log(1-pr(surv)))\)
with \(pr(surv)\) denoting the survival probability at a given time
during follow-up.\textsuperscript{49, 50} The transformed survival
probabilities are then pooled across imputed datasets and individual
time points following Rubin's rule and back-transformed via
\(1-exp(-exp(qbar))\) with \(qbar\) denoting the pooled survival
probability. The median survival time can be finally determined by
extracting the time point during follow-up at which the survival
probability drops below 0.5 for the first time.

\subsubsection{Sensitivity analyses}\label{sensitivity-analyses}

With the goal to better understand which factors could influence a
difference between RCT results and emulated trial results, a range of
sensitivity analyses will be conducted. This can comprise decisions on
the considered databases, calendar time period, covariate measurement
(e.g., measurement windows or trade-offs on sensitivity versus
specificity in measurements), approaches to missing data, selection of
covariates for imputation and propensity score models and decisions on
when to censor patients. All sensitivity analyses will be pre-specified
in the study protocol and reported using appropriate visualizations such
as forest plots.

\subsubsection{Reproducibility and consistency across
emulation}\label{reproducibility-and-consistency-across-emulation}

To ensure a transparent, reproducible and consistent way of deriving
analytic cohorts and performing statistical analyses across trial
emulations, we developed an internal R package \texttt{encore.io} with
parameterized functions. Detailed documentation can be found in the
Supplementary Material.

\newpage{}

\section{Discussion}\label{discussion}

Building on an established approach developed by the RCT DUPLICATE
project\textsuperscript{9, 30}, the ENCORE project aims to emulate 12
oncology trials using four oncology EHR data sources to inform the
potential use of RWE for regulatory science in oncology. The project
focuses on four common cancers and will evaluate the agreement of
treatment effect estimates between RCTs and their respective emulations.
Historically, administrative health claims databases have been the
backbone of most research in RWE. With increasing access to EHR data and
a maturing set of methodological approaches to draw causal inferences
from such data, the ENCORE project is one of the first of its kind to
evaluate when and how RWD can be used to deliver similar causal
conclusions compared to RCTs in the field of oncology.

The RCT DUPLICATE initiative\textsuperscript{9, 30} identified multiple
emulation challenges which we expect to also encounter in ENCORE. One
particular aspect that we may not always be able to emulate is the exact
distribution of patient characteristics of the trial population. This
can be due to the lack of data granularity and comprehensiveness to
emulate relevant eligibility criteria or the fact that large pivotal
trials in oncology are typically conducted in multiple countries
worldwide, which we will not be able to mirror given that all considered
databases reflect the US only. This may be a critical factor especially
given that the pathophysiology, prognosis and factors that drive
heterogeneous treatment outcomes of certain cancers differ between
countries. For example, some cancer indications (e.g., GI cancers) or
some genetic mutations (e.g., EGFR mutations in lung adenocarcinoma) are
much more prevalent in Asian countries compared to the US or Europe.

Another common challenge in the emulation of oncology trials is the
estimation of an ``intention-to-treat'' (ITT) analogous estimand which
is usually the primary estimand reported in oncology RCTs. Due to
intercurrent events, such as non-adherence, crossover of a high
proportion of patients from the control to the intervention arm, or
differences in subsequent therapy lines there may be estimand
differences between trial and emulation.\textsuperscript{51} Although
this is also a common challenge in the analysis of
RCTs,\textsuperscript{27} treatment protocols in routine care are often
observed to be less stringent compared to trials.\textsuperscript{8} In
order to derive comparable estimands it is therefore crucial to
understand and contextualize the proportion and timing of treatment
switching and discontinuation in both the trial and its emulation.

This issue may be even augmented with more complex treatment protocols
like combination regimens (e.g., CheckMate9LA) as compared to
monotherapies (e.g., CheckMate057) since the ascertainment of the
exposure needs to happen over a pre-defined time window which is often
difficult to calibrate in RWD. As a result, exposure misclassification
(due to overly ascertainment windows) and selection bias (due to overly
long ascertainment windows) are common trade-offs in such scenarios.
Alternative analytic approaches which target a per-protocol estimand,
such as the clone-censor-weight design,\textsuperscript{52} may be
viable options if the ITT estimand cannot be estimated due to the
aforementioned parameters, provided that necessary covariate
measurements are available to account for the artificial censoring
introduced with these methods.

\subsection{Conclusions}\label{conclusions}

RWE based on fit-for-purpose data and principled, well-designed and
reproducible studies may complement evidence coming from RCTs. Through a
systematic benchmarking approach, the ENCORE project will provide
insights as to how measurement, design and analytic decisions influence
bias and validity in oncological RWE studies.

\newpage{}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{0}{1}
\bibitem[\citeproctext]{ref-RWEFDA}
\textbf{1}. Framework for FDA's real-world evidence program (last
accessed 11/28/2024) {[}Internet{]}, 2018Available from:
\url{https://www.fda.gov/media/120060/download?attachment}

\bibitem[\citeproctext]{ref-purpura2022role}
\textbf{2}. Purpura CA, Garry EM, Honig N, et al: The role of real-world
evidence in FDA-approved new drug and biologics license applications.
Clinical Pharmacology \& Therapeutics 111:135--144, 2022

\bibitem[\citeproctext]{ref-senior2024fresh}
\textbf{3}. Senior M: Fresh from the biotech pipeline: Record-breaking
FDA approvals. Nature Biotechnology, 2024

\bibitem[\citeproctext]{ref-Merola2022}
\textbf{4}. Merola D, Schneeweiss S, Schrag D, et al: An algorithm to
predict data completeness in oncology electronic medical records for
comparative effectiveness research {[}Internet{]}. Annals of
Epidemiology 76:143--149, 2022Available from:
\url{http://dx.doi.org/10.1016/j.annepidem.2022.07.007}

\bibitem[\citeproctext]{ref-joshua2022longitudinal}
\textbf{5}. Joshua Lin K, Jin Y, Gagne J, et al: Longitudinal data
discontinuity in electronic health records and consequences for
medication effectiveness studies. Clinical Pharmacology \& Therapeutics
111:243--251, 2022

\bibitem[\citeproctext]{ref-rider2024emulations}
\textbf{6}. Rider JR, Wasserman A, Slipski L, et al: Emulations of
oncology trials using real-world data: A systematic literature review.
American journal of epidemiology kwae346, 2024

\bibitem[\citeproctext]{ref-merola2023aetion}
\textbf{7}. Merola D, Campbell U, Gautam N, et al: The aetion coalition
to advance real-world evidence through randomized controlled trial
emulation initiative: oncology. Clinical Pharmacology \& Therapeutics
113:1217--1222, 2023

\bibitem[\citeproctext]{ref-merola2024calibrating}
\textbf{8}. Merola D, Campbell U, Lenis D, et al: Calibrating
observational health record data against a randomized trial. JAMA
Network Open 7:e2436535--e2436535, 2024

\bibitem[\citeproctext]{ref-wang2023emulation}
\textbf{9}. Wang SV, Schneeweiss S, Franklin JM, et al:
\href{https://doi.org/10.1001/jama.2023.4221}{Emulation of randomized
clinical trials with nonrandomized database analyses: Results of 32
clinical trials}. Jama 329:1376--1385, 2023

\bibitem[\citeproctext]{ref-heyard2024design}
\textbf{10}. Heyard R, Held L, Schneeweiss S, et al: Design differences
and variation in results between randomised trials and non-randomised
emulations: Meta-analysis of RCT-DUPLICATE data {[}Internet{]}. BMJ
medicine 3, 2024Available from:
\url{https://doi.org/10.1136/bmjmed-2023-000709}

\bibitem[\citeproctext]{ref-encoreFDA}
\textbf{11}. Calibrating real-world evidence studies in oncology against
randomized trials: ENCORE (last accessed 11/28/2024) {[}Internet{]},
2024Available from:
\url{https://www.fda.gov/about-fda/oncology-center-excellence/calibrating-real-world-evidence-studies-oncology-against-randomized-trials-encore}

\bibitem[\citeproctext]{ref-rivera2024oncology}
\textbf{12}. Rivera DR, Eckert JC, Rodriguez-Watson C, et al: The
oncology QCARD initiative: Fostering efficient evaluation of initial
real-world data proposals. Pharmacoepidemiology and Drug Safety
33:e5818, 2024

\bibitem[\citeproctext]{ref-gatto2022structured}
\textbf{13}. Gatto NM, Campbell UB, Rubinstein E, et al: The structured
process to identify fit-for-purpose data: A data feasibility assessment
framework. Clinical Pharmacology \& Therapeutics 111:122--134, 2022

\bibitem[\citeproctext]{ref-tasneem2012database}
\textbf{14}. Tasneem A, Aberle L, Ananth H, et al: The database for
aggregate analysis of ClinicalTrials. Gov (AACT) and subsequent
regrouping by clinical specialty. PloS one 7:e33677, 2012

\bibitem[\citeproctext]{ref-eisenhauer2009new}
\textbf{15}. Eisenhauer EA, Therasse P, Bogaerts J, et al: New response
evaluation criteria in solid tumours: Revised RECIST guideline (version
1.1). European journal of cancer 45:228--247, 2009

\bibitem[\citeproctext]{ref-ThanCCR}
\textbf{16}. Ton TGN, Pal N, Trinh H, et al: Replication of overall
survival, progression-free survival, and overall response in
chemotherapy arms of nonâsmall cell lung cancer trials using real-world
data {[}Internet{]}. Clinical Cancer Research 28:2844--2853,
2022Available from: \url{https://doi.org/10.1158/1078-0432.CCR-22-0471}

\bibitem[\citeproctext]{ref-mckelvey2024evaluation}
\textbf{17}. McKelvey BA, Garrett-Mayer E, Rivera DR, et al: Evaluation
of real-world tumor response derived from electronic health record data
sources: A feasibility analysis in patients with metastatic non--small
cell lung cancer treated with chemotherapy. JCO Clinical Cancer
Informatics 8:e2400091, 2024

\bibitem[\citeproctext]{ref-rwdRECIST}
\textbf{18}. Chen L, Davis R, Lee J, et al: Comparison of response from
RECIST1.1 and abstraction in real-world patients with lung cancer.
{[}Internet{]}. Journal of Clinical Oncology 41:e21194--e21194,
2023Available from:
\url{https://ascopubs.org/doi/abs/10.1200/JCO.2023.41.16_suppl.e21194}

\bibitem[\citeproctext]{ref-rivera2022friends}
\textbf{19}. Rivera DR, Henk HJ, Garrett-Mayer E, et al: The friends of
cancer research real-world data collaboration pilot 2.0: Methodological
recommendations from oncology case studies. Clinical Pharmacology \&
Therapeutics 111:283--292, 2022

\bibitem[\citeproctext]{ref-ackerman2024measurement}
\textbf{20}. Ackerman B, Gan RW, Meyer CS, et al: Measurement error and
bias in real-world oncology endpoints when constructing external control
arms. Frontiers in Drug Safety and Regulation 4:1423493, 2024

\bibitem[\citeproctext]{ref-nikolakopoulou2014interpret}
\textbf{21}. Nikolakopoulou A, Mavridis D, Salanti G: How to interpret
meta-analysis models: Fixed effect and random effects meta-analyses. BMJ
Ment Health 17:64--64, 2014

\bibitem[\citeproctext]{ref-wang2022harmonized}
\textbf{22}. Wang SV, Pottegård A, Crown W, et al: HARmonized protocol
template to enhance reproducibility of hypothesis evaluating real-world
evidence studies on treatment effects: A good practices report of a
joint ISPE/ISPOR task force {[}Internet{]}. Value in Health
25:1663--1672, 2022Available from:
\url{https://doi.org/10.1016/j.jval.2022.09.001}

\bibitem[\citeproctext]{ref-guidelinegeneral}
\textbf{23}. ICH guideline. General principles on plan, design and
analysis of pharmacoepidemiological studies that utilize real-world data
for safety assessment of medicines M14. Available from:
Https://database.ich.org/sites/default/files/ICH\_M14\_Step3\_DraftGuideline\_2024\_0521.pdf
(last accessed 01/07/2025)

\bibitem[\citeproctext]{ref-hernan2022target}
\textbf{24}. Hernán MA, Wang W, Leaf DE: Target trial emulation: A
framework for causal inference from observational data. Jama
328:2446--2447, 2022

\bibitem[\citeproctext]{ref-hernan2016specifying}
\textbf{25}. Hernán MA, Sauer BC, Hernández-Dı́az S, et al:
\href{https://doi.org/10.1016/j.jclinepi.2016.04.014}{Specifying a
target trial prevents immortal time bias and other self-inflicted
injuries in observational analyses}. Journal of clinical epidemiology
79:70--75, 2016

\bibitem[\citeproctext]{ref-meyer2020open}
\textbf{26}. Meyer A-M, Davies J, Taylor M, et al: Open cohorts and
ghost-time bias in real world data, in PHARMACOEPIDEMIOLOGY AND DRUG
SAFETY. WILEY 111 RIVER ST, HOBOKEN 07030-5774, NJ USA, 2020, pp
426--426

\bibitem[\citeproctext]{ref-rufibach2018}
\textbf{27}. Rufibach K: Treatment effect quantification for
time{-}to{-}event endpoints{\textendash}Estimands, analysis strategies,
and beyond {[}Internet{]}. Pharmaceutical Statistics 18:145--165,
2018Available from: \url{http://dx.doi.org/10.1002/pst.1917}

\bibitem[\citeproctext]{ref-hernan2022causal}
\textbf{28}. Hernán MA: Causal analyses of existing databases: No power
calculations required. Journal of clinical epidemiology 144:203--205,
2022

\bibitem[\citeproctext]{ref-schoenfeld1981asymptotic}
\textbf{29}. Schoenfeld D: The asymptotic properties of nonparametric
tests for comparing survival distributions. Biometrika 68:316--319, 1981

\bibitem[\citeproctext]{ref-franklin2020nonrandomized}
\textbf{30}. Franklin JM, Pawar A, Martin D, et al: Nonrandomized
real-world evidence to support regulatory decision making: Process for a
randomized trial replication project. Clinical Pharmacology \&
Therapeutics 107:817--826, 2020

\bibitem[\citeproctext]{ref-schneeweiss2019graphical}
\textbf{31}. Schneeweiss S, Rassen JA, Brown JS, et al: Graphical
depiction of longitudinal study designs in health care databases. Annals
of internal medicine 170:398--406, 2019

\bibitem[\citeproctext]{ref-rubin1976inference}
\textbf{32}. Rubin DB: Inference and missing data. Biometrika
63:581--592, 1976

\bibitem[\citeproctext]{ref-weberpals2024smdi}
\textbf{33}. Weberpals J, Raman SR, Shaw PA, et al: Smdi: An r package
to perform structural missing data investigations on partially observed
confounders in real-world evidence studies. JAMIA open 7:ooae008, 2024

\bibitem[\citeproctext]{ref-weberpals2024}
\textbf{34}. Weberpals J, Raman SR, Shaw PA, et al: A principled
approach to characterize and analyze partially observed confounder data
from electronic health records {[}Internet{]}. Clinical Epidemiology
16:329--343, 2024Available
from:\href{\%20\%0A\%20\%20\%20\%20\%0A\%20\%20\%20\%20\%0A\%20\%20\%20\%20\%20\%20\%20\%20https://www.tandfonline.com/doi/abs/10.2147/CLEP.S436131\%0A\%20\%20\%20\%20\%0A\%0A}{
https://www.tandfonline.com/doi/abs/10.2147/CLEP.S436131 }

\bibitem[\citeproctext]{ref-moreno2018canonical}
\textbf{35}. Moreno-Betancur M, Lee KJ, Leacy FP, et al: Canonical
causal diagrams to guide the treatment of missing data in epidemiologic
studies. American journal of epidemiology 187:2705--2715, 2018

\bibitem[\citeproctext]{ref-tompsett2018use}
\textbf{36}. Tompsett DM, Leacy F, Moreno-Betancur M, et al: On the use
of the not-at-random fully conditional specification (NARFCS) procedure
in practice. Statistics in medicine 37:2338--2353, 2018

\bibitem[\citeproctext]{ref-weberpals2024hdmi}
\textbf{37}. Weberpals J, Shaw PA, Lin KJ, et al: High-dimensional
multiple imputation (HDMI) for partially observed confounders including
natural language processing-derived auxiliary covariates {[}Internet{]}.
American Journal of Epidemiology, 2025Available from:
\url{https://arxiv.org/abs/2405.10925}

\bibitem[\citeproctext]{ref-Shaw2024}
\textbf{38}. Shaw CK P., Williamson BD: Assessing treatment effects in
observational data with missing confounders: A comparative study of
practical doubly-robust and traditional missing data methods
{[}Internet{]}. GitHub repository, 2024Available from:
\url{https://github.com/PamelaShaw/Missing-Confounders-Methods}

\bibitem[\citeproctext]{ref-cox1972regression}
\textbf{39}. Cox DR: Regression models and life-tables. Journal of the
royal statistical society. Series B (Methodological) 34:187--220, 1972

\bibitem[\citeproctext]{ref-kaplan1958nonparametric}
\textbf{40}. Kaplan EL, Meier P: Nonparametric estimation from
incomplete observations. Journal of the American statistical association
53:457--481, 1958

\bibitem[\citeproctext]{ref-becker2020enhanced}
\textbf{41}. Becker T, Weberpals J, Jegg A, et al: An enhanced
prognostic score for overall survival of patients with cancer derived
from a large real-world cohort. Annals of Oncology 31:1561--1568, 2020

\bibitem[\citeproctext]{ref-leyrat2019propensity}
\textbf{42}. Leyrat C, Seaman SR, White IR, et al: Propensity score
analysis with partially observed covariates: How should multiple
imputation be used? Statistical methods in medical research 28:3--19,
2019

\bibitem[\citeproctext]{ref-pishgar2020matchthem}
\textbf{43}. Pishgar F, Greifer N, Leyrat C, et al: MatchThem:: Matching
and weighting after multiple imputation. arXiv preprint arXiv:200911772,
2020

\bibitem[\citeproctext]{ref-rubin2018multiple}
\textbf{44}. Rubin DB: Multiple imputation, in Flexible imputation of
missing data, second edition. Chapman; Hall/CRC, 2018, pp 29--62

\bibitem[\citeproctext]{ref-van2011mice}
\textbf{45}. Van Buuren S, Groothuis-Oudshoorn K: Mice: Multivariate
imputation by chained equations in r. Journal of statistical software
45:1--67, 2011

\bibitem[\citeproctext]{ref-austin2009balance}
\textbf{46}. Austin PC: Balance diagnostics for comparing the
distribution of baseline covariates between treatment groups in
propensity-score matched samples. Statistics in medicine 28:3083--3107,
2009

\bibitem[\citeproctext]{ref-franklin2014metrics}
\textbf{47}. Franklin JM, Rassen JA, Ackermann D, et al: Metrics for
covariate balance in cohort studies of causal effects. Statistics in
medicine 33:1685--1699, 2014

\bibitem[\citeproctext]{ref-stuart2013prognostic}
\textbf{48}. Stuart EA, Lee BK, Leacy FP: Prognostic score--based
balance measures can be a useful diagnostic for propensity score methods
in comparative effectiveness research. Journal of clinical epidemiology
66:S84--S90, 2013

\bibitem[\citeproctext]{ref-warwick28685}
\textbf{49}. Marshall A(Andrea), Billingham LJ, Bryan S: Can we afford
to ignore missing data in cost-effectiveness analyses? {[}Internet{]}.
European Journal of Health Economics Vol.10:1--3, 2009Available from:
\url{http://dx.doi.org/10.1007/s10198-008-0129-y}

\bibitem[\citeproctext]{ref-morisot2015prostate}
\textbf{50}. Morisot A, Bessaoud F, Landais P, et al: Prostate cancer:
Net survival and cause-specific survival rates after multiple
imputation. BMC medical research methodology 15:1--14, 2015

\bibitem[\citeproctext]{ref-manitz2022estimands}
\textbf{51}. Manitz J, Kan-Dobrosky N, Buchner H, et al: Estimands for
overall survival in clinical trials with treatment switching in
oncology. Pharmaceutical Statistics 21:150--162, 2022

\bibitem[\citeproctext]{ref-gaber2024mystifying}
\textbf{52}. Gaber CE, Ghazarian AA, Strassle PD, et al: De-mystifying
the clone-censor-weight method for causal research using observational
data: A primer for cancer researchers. Cancer Medicine 13:e70461, 2024

\end{CSLReferences}

\newpage{}

\section*{Tables}\label{tables}
\addcontentsline{toc}{section}{Tables}

\begin{table}[h]

\caption{\label{tbl-criteria}Criteria to select eligible trials for
emulation in ENCORE.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../tables/Table_1_trial_eligibility.png}}

}

\end{table}%

\newpage{}

\begin{table}[h]

\caption{\label{tbl-rcts}Tentative list of randomized controlled trials
(RCTs) considered for emulation.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../tables/Table_2_trial_selection.png}}

}

\end{table}%

\newpage{}

\begin{table}[h]

\caption{\label{tbl-metrics}Visualization of agreement metrics example.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../tables/Table_3_agreement_metrics.png}}

}

\end{table}%

\newpage{}

\section*{Figures}\label{figures}
\addcontentsline{toc}{section}{Figures}

\begin{figure}[h]

\caption{\label{fig-process}Systematic process to understand
effectiveness claims of oncology trials using real-world evidence.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{../figures/process.png}}

}

\end{figure}%

\newpage{}

\begin{figure}[h]

\caption{\label{fig-initiators}Example visualization of descriptive drug
utilization analyses displaying a) initiation trends between compared
regimens based on calendar time, b) cumulative rate of patients
switching to another line of treatment.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{Figure_2_utlization.png}}

}

\end{figure}%

\newpage{}

\begin{figure}[h]

\caption{\label{fig-balance}Assessment of a) covariate balance and b)
distributional balance of a prognostic score for overall survival before
and after propensity score matching or weighting across multiple imputed
datasets.}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{Figure_3_balance.png}}

}

\end{figure}%

\end{document}