### 2025-05-28 manuscript/supplement.qmd
additions on runner up clinical trial candidates for emulation

---
title: "Supplementary Material"
subtitle: "A Process for the Emulation of Comparative Oncology Trials with Real-world Evidence (ENCORE)"

echo: false
number-sections: false
toc: true

format: 
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \usepackage{pdfpages}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
    fig-cap-location: top
    crossref:
      fig-title: '**Supplementary Figure**'
      tbl-title: '**Supplementary Table**'

editor: visual
bibliography: references.bib
#csl: american-journal-of-epidemiology.csl
---

```{r}
#| label: setup_supplement
#| include: false

library(here)
library(knitr)
library(readxl)
library(gt)
library(dplyr)

# table_cap <- captioner(prefix = "Supplementary Table")
```

{{< pagebreak >}}

# Supplementary Figures and Tables

```{r}
#| label: fig-consort-nsclc
#| fig-cap: "CONSORT diagram for non-small cell lung cancer (NSCLC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_nsclc.png"))
```

*Note that Supplementary @fig-consort-nsclc results in four shortlisted candidates because both CheckMate017/057 have been both shortlisted and only differ in the squamous versus nonsquamous histological eligibility of the trial population.*

{{< pagebreak >}}

```{r}
#| label: fig-consort-breast
#| fig-cap: "CONSORT diagram for breast cancer (BC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_bc.png"))
```

{{< pagebreak >}}

```{r}
#| label: fig-consort-crc
#| fig-cap: "CONSORT diagram for colorectal cancer (CRC) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_crc.png"))
```

{{< pagebreak >}}

```{r}
#| label: fig-consort-mm
#| fig-cap: "CONSORT diagram for multiple myeloma (MM) top candidate trials."
#| out-width: "115%"

include_graphics(here("figures", "consort_mm.png"))
```

```{r}
#| label: tbl-runner-ups
#| tbl-cap: "Oncology trial emulation potential runner up candidates."
#| tbl-cap-location: top

# read excel blueprint
runner_ups <- read_excel(
  path = here("tables", "helper_tables.xlsx"),
  sheet = "runner_ups"
  )

# convert to gt table
runner_ups_gt <- runner_ups |> 
  group_by(cancer) |> 
  gt() |> 
  cols_label(
    nctid = md("**NCTID**"),
    acronym = md("**Acronym**"),
    setting = md("**Clinical setting**"),
    line = md("**Line of therapy**"),
    exposures = md("**Treatment comparison**")
    ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
    )

# save image to display within manuscript document
suppressMessages(runner_ups_gt |> 
  gtsave(
    filename = here("tables", "STable_runner_ups.png"),
    expand = 10, # amount of whitespace,
    zoom = 5 # higher resolution
    ))

include_graphics(path = here("tables", "STable_runner_ups.png"))
```

{{< pagebreak >}}

# R package documentation

The following documentation for the internal `encore.io` packages describes and details reproducible functions to query analytic cohorts consistently across trial emulations.

::: callout-important
The `encore.io` package will be continually developed throughout the ENCORE project and the following documentation is a version-controlled snapshot at the time of this publication. Updated versions will be published with the corresponding protocols for each trial emulation separately.
:::

\includepdf[pages=-]{encore.io_0.2.0.pdf}

{{< pagebreak >}}

$\vspace{0.25cm}$

*Supplementary Material last updated: `r Sys.Date()`*